---
title: "Predicting Labor.RMD"
author: "Russ Conte"
date: "12/27/2021"
output: html_document
---

This is an analysis of labor data to predict future labor trends. The majority of the data is from the United States Department of Labor Statistics.

Outline:

1. Retreive the labor reports
2. Combine the reports into one time series file (called a tsibble in this system)
3. Run time series models to predict future numbers
4. Run time series models to predict future changes

Part 1: Retreive the data

All employees: https://beta.bls.gov/dataViewer/view/timeseries/CES0000000001
Consumer Price Index: https://beta.bls.gov/dataViewer/view/timeseries/CUSR0000SA0
Unemployment: 	https://beta.bls.gov/dataViewer/view/timeseries/LNS14000000
Average Hourly Earnings: 	https://beta.bls.gov/dataViewer/view/timeseries/CES0500000003
Worker Productivity: 	https://beta.bls.gov/dataViewer/view/timeseries/PRS85006092 # Quarterly
Temporary help: https://beta.bls.gov/dataViewer/view/timeseries/CES6056132001

Turnover Level: https://data.bls.gov/timeseries/JTS000000000000000TSL
Turnover Rate: https://data.bls.gov/timeseries/JTS000000000000000TSR
Hires Level: https://beta.bls.gov/dataViewer/view/timeseries/JTS000000000000000HIL
Hires rate: https://beta.bls.gov/dataViewer/view/timeseries/JTS000000000000000HIR
Job Openings Level: https://beta.bls.gov/dataViewer/view/timeseries/JTS000000000000000JOL
Job Openings Rate: https://beta.bls.gov/dataViewer/view/timeseries/JTS000000000000000JOR
Layoffs and Discharges Level: https://beta.bls.gov/dataViewer/view/timeseries/JTS000000000000000LDL
Layoffs and Discharges Rate: https://beta.bls.gov/dataViewer/view/timeseries/JTS000000000000000LDR
Other Separations Level: https://beta.bls.gov/dataViewer/view/timeseries/JTS000000000000000OSL
Other Separations Rate: https://beta.bls.gov/dataViewer/view/timeseries/JTS000000000000000OSR
Quits Level: https://beta.bls.gov/dataViewer/view/timeseries/JTS000000000000000QUL
Quits Rate: https://beta.bls.gov/dataViewer/view/timeseries/JTS000000000000000QUR
Total Separations Level: https://beta.bls.gov/dataViewer/view/timeseries/JTS000000000000000TSL
Total Separations Rate: https://beta.bls.gov/dataViewer/view/timeseries/JTS000000000000000TSR

```{r Retreive the data}

library(tidyverse)
library(fpp3)
library(clock)

#Convert files to tsibble format
All_employees <- All_employees %>% 
  mutate(Month = yearmonth(Label), Total_employees = Value) %>% 
  as_tsibble(index = Month) %>% 
  mutate(Total_employees_diff = difference(Value)) %>% 
  select(Total_employees:Total_employees_diff)
Total_employees <- All_employees$Total_employees
index = All_employees$Month

Consumer_price_index <- Consumer_price_index %>% 
  mutate(CPI = Value, Month = yearmonth(Label)) %>% 
  as_tsibble(index = Month) %>% 
  mutate(CPI_diff = difference(Value)) %>% 
  select(CPI:CPI_diff)

Unemployment <- Unemployment %>% 
  mutate(Month = yearmonth(Label), Num_unemployed = Value) %>% 
  as_tsibble(index = Month) %>% 
  mutate(Unemployment_diff = difference(Value)) %>% 
  select(Num_unemployed:Unemployment_diff)

Average_hourly_earnings <- Average_hourly_earnings %>%
  mutate(Avg_hourly_earning = Value, Month = yearmonth(Label)) %>%
  as_tsibble(index = Month) %>%
  mutate(Avg_earnings_diff = difference(Value)) %>%
  select(Avg_hourly_earning:Avg_earnings_diff)

Temp_labor <- Temp_labor %>% 
  mutate(Month = yearmonth(Label), Num_temps = Value) %>% 
  as_tsibble(index = Month) %>% 
  mutate(temp_labor_diff = difference(Value)) %>% 
  select(Num_temps:temp_labor_diff)

Turnover_rate <- Turnover_rate %>%
  pivot_longer(cols = Jan:Dec)
Turnover_rate
Turnover_rate$name = match(Turnover_rate$name, month.abb)
Turnover_rate <- Turnover_rate %>%
    mutate(date = date_build(year = Year, month = name),
          Month = yearmonth(date),
         key = 0,
         turnover_rate = value,
         turnover_rate_diff = difference(turnover_rate)) %>%
  select(Month, key, turnover_rate, turnover_rate_diff)
Turnover_rate <- Turnover_rate %>%
  as_tsibble(index = Month, key = key)

Hires <- Hires %>% 
  mutate(Month = yearmonth(Label), key = 0, Hires = Value) %>% 
  as_tsibble(index = Month, key = key) %>% 
  mutate(Hires_diff = difference(Value)) %>% 
  select(Month:Hires_diff)

Job_openings <- Job_openings %>% 
  mutate(Month = yearmonth(Label), key = 0, Job_openings = Value) %>% 
  as_tsibble(index = Month, key = key) %>% 
  mutate(Job_openings_diff = difference(Value)) %>% 
  select(Month:Job_openings_diff)

Layoffs_and_discharges <- Layoffs_and_discharges %>% 
  mutate(Month = yearmonth(Label), key = 0, Layoffs_and_discharges = Value) %>% 
  as_tsibble(index = Month, key = key) %>% 
  mutate(Layoffs_and_discharges_diff = difference(Value)) %>% 
  select(Month:Layoffs_and_discharges_diff)

Other_separations <- Other_separations %>% 
  mutate(Month = yearmonth(Label), key = 0, Other_separations = Value) %>% 
  as_tsibble(index = Month, key = key) %>% 
  mutate(Other_separations_diff = difference(Value)) %>% 
  select(Month:Other_separations_diff)

Quits <- Quits %>% 
  mutate(Month = yearmonth(Label), key = 0, Quits = Value) %>% 
  as_tsibble(index = Month, key = key) %>% 
  mutate(Quits_diff = difference(Value)) %>% 
  select(Month:Quits_diff)

Total_Separations <- Total_Separations %>% 
  mutate(Month = yearmonth(Label), key = 0, Total_separations = Value) %>% 
  as_tsibble(index = Month, key = key) %>% 
  mutate(Total_separations_diff = difference(Value)) %>% 
  select(Month:Total_separations_diff)

key = 0

```

Make a tsibble (time series) that combines all the data into one file

```{r Combine all the features into one file that will be used to make predictions}

labor <- 
  tibble(
    Total_Employees = Total_employees,
    CPI = Consumer_price_index$CPI,
    Unemployment_rate = Unemployment$Num_unemployed,
    Num_temporary_employees = Temp_labor$Num_temps,
    Turnover = Turnover_rate$turnover_rate[1:length(All_employees$Month)],
    Num_hires = Hires$Hires[1:length(All_employees$Month)],
    Num_job_openings = Job_openings$Job_openings[1:length(All_employees$Month)],
    Num_layoffs_and_discharges = Layoffs_and_discharges$Layoffs_and_discharges[1:length(All_employees$Month)],
    Num_other_separations = Other_separations$Other_separations[1:length(All_employees$Month)],
    Num_quits = Quits$Quits[1:length(All_employees$Month)],
    Num_total_separations = Total_Separations$Total_separations[1:length(All_employees$Month)]
    ) %>% 
  mutate(Month = All_employees$Month) %>% 
  as_tsibble(index = Month)

labor

```

